---
title: "Chulwalar Case Study"
author: "Aravind Veluchamy & Sandhya Amaresh"
date: "July 27, 2016"
output:
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Chulwalar case study - In this case study, we analyse the data for Blue Etel flower and build a forecast model based on the timeseries data for the same

Steps:
1 - Data extraction and clean up
2 - Exploratory data analysis
3 - Correration 
4 - Forecast Model , comparision
5 - Results

### 1 - Data extraction and clean up
#### Read the As Is, Planned & Indicators files from the DataFiles directory into a dataframe
#### Spilt and extract the data for Blue Etel from the AsIs and Plan Dataframes 
#### Transform the data to a timeseries object
```{r}
setwd("C:/Users/samaresh/SMU/Doing Data Science/ChulwalarCaseStudy")
ImportedAsIsData <- read.csv("DataFiles/ImportedAsIsDataChulwalar.csv", header = F, sep=";", fill = T) 
ImportedPlanData <- read.csv("DataFiles/ImportedPlanDataChulwalar.csv", header = F, sep=";", fill = T) 
ImportedIndicators <- read.csv("DataFiles/ImportedIndicatorsChulwalar.csv", header = F, sep=";", fill = T) 


BlueEtelAsIsVector <- c(ImportedAsIsData [58:69,2],ImportedAsIsData [58:69,3],ImportedAsIsData [58:69,4],ImportedAsIsData [58:69,5],ImportedAsIsData [58:69,6],ImportedAsIsData [58:69,7])
BlueEtelPlanVector <- c(ImportedPlanData[58:69,2],ImportedPlanData[58:69,3],ImportedPlanData[58:69,4],ImportedPlanData[58:69,5],ImportedPlanData[58:69,6],ImportedPlanData[58:69,7])
RedEtelPlanVector <- c(ImportedPlanData[72:83,2],ImportedPlanData[72:83,3],ImportedPlanData[72:83,4],ImportedPlanData[72:83,5],ImportedPlanData[72:83,6],ImportedPlanData[72:83,7])


BlueEtelAsIs <- ts(BlueEtelAsIsVector, start=c(2008,1), end=c(2013,12), frequency=12)
BlueEtelPlan <- ts(BlueEtelPlanVector, start=c(2008,1), end=c(2013,12), frequency=12)
```

### 2 - Exploratory Data Analysis
#### Plot the ts for AsIs and Planned Data 
#### Obtain a correlation b/w the AsIs and Planned
#### Decompose the TS using STL to analyse the seasonal, trend and irregular components

```{r}
library(fpp)
options(scipen = 5)
par(mfrow=c(1,2))
plot(BlueEtelAsIs, col="orange", main="BlueEtelAsIs")
plot(BlueEtelPlan, col="orange", main="BlueEtelPlan")
```
```{r}
cor(BlueEtelAsIs , BlueEtelPlan)

# The results show a very high planning accuracy. 
BlueEltelAsIs_lm <- lm(BlueEtelAsIs ~ BlueEtelPlan , data = BlueEtelAsIs)
summary(BlueEltelAsIs_lm)

BlueEtelAsIs_tslm <- tslm(BlueEtelAsIs ~ BlueEtelPlan )
summary(BlueEltelAsIs_lm)

BlueEtelAsIs_stl <- stl(BlueEtelAsIs , s.window=5)

par(mfrow=c(3,2))

plot(BlueEtelAsIs_stl, col="black", main="BlueEtelAsIs_stl")

par(mfrow=c(1,2))

plot(BlueEtelAsIs_stl$time.series[,"trend"], col="orange")

monthplot(BlueEtelAsIs_stl$time.series[,"seasonal"], main="", ylab="Seasonal")
```

#### Based on the above plots we observer the following -
#### The STL decompose the BlueEtelAsIs data into 4 plots - the Observed, Seasonal, Trend and Random
#### From the Trend plot we can see a -ve trend in the export of Blue Etel from 2008 to 2011, and the trend becomes positive from 2011 thru 2013
#### From the seasonal plot we can see that the exports fall during summer and then after the exquinox around Sept the exports pick up and we see a small reduction in exports during december which coincides with the indpendence day celebrations and again rises till march equinox



### 4 - Forecast Models and Comparisons
#### Model 1 - Simple exponential smoothing
#### In this model the past values are used to calculate a forecast. The strength with which each value influences the forecast is weakened with help of a smoothing parameter. Thus we are dealing with a weighted average, whose values fade out the longer ago they were in the past.
```{r}
Model_ses <- ses(BlueEtelAsIs, h=12)
summary(Model_ses)
plot(Model_ses)
```
#### The Akaike's Information Criterion(AIC/AICc) or the Bayesian Information 
#### Criterion (BIC) should be at minimum.
```{r}
plot(Model_ses, plot.conf=FALSE, ylab="Exports Chulwalar  )", xlab="Year", main="", fcol="white", type="o")
lines(fitted(Model_ses), col="green", type="o")
lines(Model_ses$mean, col="blue", type="o")
legend("topleft",lty=1, col=c(1,"green"), c("data", expression(alpha == 0.671)),pch=1)
```

#### Model 2 - HOlt's Linear Method
#### This model in addition to SES model uses the trend as well. When neither alpha nor beta are provided, they paramets are optimised
```{r}
Model_holt_1 <- holt(BlueEtelAsIs,h=12)
summary(Model_holt_1)
plot(Model_holt_1)
```

####  The trend is exponential if the intercepts(level) and the gradient (slope) are
####  multiplied with eachother. The values are worse. As the Beta was very low in 
####  the optimisation, the forecast is very similar to the ses() model. 

```{r}
Model_holt_2<- holt(BlueEtelAsIs, exponential=TRUE,h=12)
summary(Model_holt_2)
plot(Model_holt_2)
```
####  As such simple trends tend to forecast the future to positively, we have added a dampener.
#### Similar values to that of Model_holt_1 
```{r}
Model_holt_3 <- holt(BlueEtelAsIs, damped=TRUE,h=12)
summary(Model_holt_3)
plot(Model_holt_3)
```
####  This also works for exponential trends. 
####  The values remain worse. 
```{r}
Model_holt_4 <- holt(BlueEtelAsIs, exponential=TRUE, damped=TRUE,h=12)
summary(Model_holt_4)
plot(Model_holt_4)
```

####  level and slope can be plotted individually for each model.
```{r}
plot(Model_holt_1$model$state)
plot(Model_holt_2$model$state)
plot(Model_holt_3$model$state)
plot(Model_holt_4$model$state)

plot(Model_holt_1, plot.conf=FALSE, ylab="Exports Chulwalar  )", xlab="Year", main="", fcol="white", type="o")
lines(fitted(Model_ses), col="purple", type="o")
lines(fitted(Model_holt_1), col="blue", type="o")
lines(fitted(Model_holt_2), col="red", type="o")
lines(fitted(Model_holt_3), col="green", type="o")
lines(fitted(Model_holt_4), col="orange", type="o")
lines(Model_ses$mean, col="purple", type="o")
lines(Model_holt_1$mean, col="blue", type="o")
lines(Model_holt_2$mean, col="red", type="o")
lines(Model_holt_3$mean, col="green", type="o")
lines(Model_holt_4$mean, col="orange", type="o")
legend("topleft",lty=1, col=c(1,"purple","blue","red","green","orange"), c("data", "SES","Holts auto", "Exponential", "Additive Damped", "Multiplicative Damped"),pch=1)
```

####  As these forecasts are not very convincing at the moment, there is no need 
####  to export the data.

#### Model 3 - Holt - Winter's seasonality model
#### This model expands the Holt's model to futher include the seasonality aspect
#### There are 2 models under this, the additive model and multiplicative model
#### 3.1 Additive Model
```{r}
Model_hw_1 <- hw(BlueEtelAsIs ,seasonal="additive",h=12)
summary(Model_hw_1)
plot(Model_hw_1)
```

#### 3.2 Multiplicative Model
```{r}
Model_hw_2 <- hw(BlueEtelAsIs ,seasonal="multiplicative",h=12)
summary(Model_hw_2)
plot(Model_hw_2)
```

```{r}
plot(Model_hw_1, ylab="Exports Chulwalar  ", plot.conf=FALSE, type="o", fcol="white", xlab="Year")
lines(fitted(Model_hw_1), col="red", lty=2)
lines(fitted(Model_hw_2), col="green", lty=2)
lines(Model_hw_1$mean, type="o", col="red")
lines(Model_hw_2$mean, type="o", col="green")
legend("topleft",lty=1, pch=1, col=1:3, c("data","Holt Winters' Additive","Holt Winters' Multiplicative"))

# In order to use the results later, they need to be converted into point forcasts.
Model_hw_1_df <-as.data.frame(Model_hw_1) 
Model_hw_1_PointForecast <- ts(Model_hw_1_df$"Point Forecast", start=c(2014,1), end=c(2014,12), frequency=12)
Model_hw_1_PointForecast
Model_hw_2_df <-as.data.frame(Model_hw_2) 
Model_hw_2_PointForecast <- ts(Model_hw_2_df$"Point Forecast", start=c(2014,1), end=c(2014,12), frequency=12)
Model_hw_2_PointForecast

# Output instruction for the data export of the results for further use in Excel.
#write.csv(Model_hw_1_PointForecast,file='Model_hw_1_PointForecast.csv')
#write.csv(Model_hw_2_PointForecast,file='Model_hw_2_PointForecast.csv')
```

### Results -
### Based on the results, we see that the multiplicative model is slightly better than the additive model.
### The above table shows the forecast numbers based on both the models
